The sound of teaching music: Expert pianists' performance modulations for novices

### 2022-09-01 paper/manuscript.tex
finalise the manuscript with some minor corrections
Signed-off-by: Atsuko Tominaga <Tominaga_Atsuko@phd.ceu.edu>

In the teaching condition, participants were instructed to play the excerpt of music as if they were teaching it to students. It was mentioned that the students already knew the sequence of the tones and that they were trying to learn how to perform the piece with the notated expression by listening to the participant's performance. In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience (see details in \emph{\protect\hyperlink{supplementary}{Supplementary material}}). Participants played the piece 8 times per technique per condition, so there were 32 trials in total (2 conditions x 2 techniques x 8 trials). The order of the conditions was blocked and counterbalanced across participants. The order of the techniques within each condition was also blocked and counterbalanced across participants. A leading metronome (80 quarter beats per minute, 8 beats) indicated the target tempo before each trial.

@@

A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Dynamics (forte vs.~piano) showed that there was a significant main effect of Dynamics (\emph{F}(1,30) = 369, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.74) and a significant interaction between Condition and Dynamics (\emph{F}(1,30) = 6.83, \emph{p} = 0.014, \(\eta_G^2\) = 0.003, \emph{Fig \ref{fig:vel-1}}). Post-hoc comparisons based on the estimated marginal means with Tukey adjustment revealed that KV was higher in the teaching condition {[}\emph{M} = 82.78, \emph{SD} = 8.31{]} than in the performing condition {[}\emph{M} = 81.65, \emph{SD} = 7.43{]} (\emph{p} = 0.068) when instructed to play forte. There was no significant difference in KV between the teaching {[}\emph{M} = 60.63, \emph{SD} = 5.08{]} and performing condition {[}\emph{M} = 60.99, \emph{SD} = 4.41{]} (\emph{p} = 0.24) when instructed to play piano.

@@

We employed the same procedure as in Experiment 1 with several modifications. First, all participants were required to memorise the piece without expressive notations (i.e., \emph{Fig \ref{fig:stimuli-2} A}) prior to the experiment so that they had enough time to familiarise themselves with the piece and perform it without pitch errors while performing it with different notated expressions in the lab. Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations because there had been less emphasis on expression in the instruction in the performing condition in Experiment 1 (see details in \emph{\protect\hyperlink{supplementary}{Supplementary material}}). Third, participants could choose their preferred tempo from one of three options (100, 110 and 120 quarter beats per minute) because some evidence show that musicians have their own preferred tempo, which affects their temporal variability\textsuperscript{19}. The chosen tempo was again cued by a leading metronome. To make sure that participants memorised the piece and had sufficient motor skills, we asked participants to perform the piece without looking at the original sheet music (\emph{Fig \ref{fig:stimuli-2} A}). Those who could not perform the piece without pitch errors twice consecutively within five attempts were excluded from data analysis (as a result, one participant was excluded). The rest of the procedure was identical to Experiment 1.

@@

One limitation of the study is that although we instructed participants to imagine a situation in which they were teaching musical expression to students, there was no feedback from actual students. It can be expected that teachers would also make didactic modulations in some domains other than acoustic properties, such as gestures if they are physically present in front of students. Moreover, given that teachers modulate their demonstration throughout ongoing interactions with learners\textsuperscript{20}, future studies are needed to investigate how experts dynamically adapt their performance to their students' skill level and demonstrated abilities. Also, some of the participants in the current study did not have teaching experience at all and even those who had teaching experience had taught for only a few years. It would be important to investigate if our findings would be observed in participants with extensive experience teaching the piano.

@@

Musicians and music teachers tend to consider expressive skills as performers' most important skills\textsuperscript{21}. In the present study, experts focused on the teaching of expressions that were notated. However, expression also has other facets\textsuperscript{22} that are not only piece-related (e.g., notated expressions) but performer- or context-related (e.g., performer's interpretation, listening contexts such as recordings or concerts). The ultimate goal of expressive performance is considered to lie in developing performers' own interpretations of music and convey affects and emotions by using expressive tools such as articulation, dynamics, tempo and timing\textsuperscript{23}. The current research opens up the possibility of studying the teaching of expressive performance by demonstrating that pianists could signal didactic intentions without relying on verbal instructions, which are heavily used in actually teaching settings\textsuperscript{e.g., 15}. It would be interesting to examine how teachers play (i.e., demonstrate) relates to verbal instructions they give.

The present findings extend earlier research on teaching-related action modulations. First, our study sheds light on the teaching of technical skills that are an integral part of skill acquisition in artistic domains. We showed that compared to an expressive performance baseline, experts made specific modulations to teach particular techniques. Second, the specificity of the observed modulations supports the idea that teaching comprises more than generic modulations like slowing down or overall exaggeration that may draw learners' attention. Rather, expert demonstrators seem to follow principles of relevance in communication\textsuperscript{24}, highlighting only those aspects they are intending to demonstrate. How learners benefit from the perceptual and motor cues that come with specific exaggerations, and whether understanding the teacher's intentions explicitly adds to the learning success are important questions for future research.

@@

\leavevmode\vadjust pre{\hypertarget{ref-zamm_2018}{}}%
\CSLLeftMargin{19. }%
\CSLRightInline{Zamm, A., Wang, Y. \& Palmer, C. \href{https://doi.org/10.1177/0748730418783651}{Musicians' {Natural Frequencies} of {Performance Display Optimal Temporal Stability}}. \emph{J Biol Rhythms} \textbf{33}, 432--440 (2018).}

\leavevmode\vadjust pre{\hypertarget{ref-okazaki_2019}{}}%
\CSLLeftMargin{20. }%
\CSLRightInline{Okazaki, S., Muraoka, Y. \& Osu, R. \href{https://doi.org/10.1038/s41598-019-44049-x}{Teacher-learner interaction quantifies scaffolding behaviour in imitation learning}. \emph{Sci Rep} \textbf{9}, 1--13 (2019).}

@@

\leavevmode\vadjust pre{\hypertarget{ref-sperber_1995}{}}%
\CSLLeftMargin{24. }%
\CSLRightInline{Sperber, D. \& Wilson, D. \emph{Relevance: {Communication} and cognition, 2nd ed}. ({Blackwell Publishing}, 1995).}

@@

\hypertarget{supplementary}{%
\section{Supplementary material}\label{supplementary}}

@@

\hypertarget{temporal-variability}{%
\section{2. Temporal Variability}\label{temporal-variability}}

@@

\hypertarget{temporal-grouping}{%
\section{3. Temporal grouping}\label{temporal-grouping}}

@@

\hypertarget{additional-analysis-for-dynamics-difference}{%
\section{4. Additional analysis for dynamics difference}\label{additional-analysis-for-dynamics-difference}}

@@

\hypertarget{descriptive-statistics-depending-on-participants-teaching-experience}{%
\section{5. Descriptive statistics depending on participants' teaching experience}\label{descriptive-statistics-depending-on-participants-teaching-experience}}

\caption{\label{tab:ioi-teaching-desc-1}Experiment 1:}

\begin{tabular}{lllllllll}
\toprule
Condition & \multicolumn{1}{c}{Technique} & \multicolumn{1}{c}{Teaching} & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c}{SEM} & \multicolumn{1}{c}{Median} & \multicolumn{1}{c}{IQR}\\
\midrule
Performing & Articulation & Yes & 10 & 181.27 & 10.99 & 3.47 & 177.37 & 14.91\\
Teaching & Articulation & Yes & 10 & 184.72 & 13.02 & 4.12 & 179.21 & 16.76\\
Performing & Articulation & No & 21 & 193.49 & 18.75 & 4.09 & 185.83 & 26.43\\
Teaching & Articulation & No & 21 & 196.67 & 20.52 & 4.48 & 192.13 & 32.33\\
Performing & Dynamics & Yes & 10 & 182.43 & 11.52 & 3.64 & 181.67 & 13.07\\
Teaching & Dynamics & Yes & 10 & 187.15 & 11.73 & 3.71 & 187.91 & 15.67\\
Performing & Dynamics & No & 21 & 191.48 & 18.56 & 4.05 & 188.17 & 23.11\\
Teaching & Dynamics & No & 21 & 193.50 & 21.22 & 4.63 & 186.28 & 25.29\\
\bottomrule
\addlinespace
\end{tabular}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} Note}
\end{tablenotes}

\caption{\label{tab:art-teaching-desc-1}Experiment 1:}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} Note}
\end{tablenotes}

\caption{\label{tab:dyn-teaching-desc-1}Experiment 1:}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} Note}
\end{tablenotes}

\caption{\label{tab:dyn-diff-teaching-desc-1}Experiment 1:}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} Note}
\end{tablenotes}

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-ioi-teaching-2-1} \caption{\label{fig:ioi-teaching-2}Experiment 2: Comparison between those with and without piano teaching experience in terms of IOIs (ms) when playing the piece with either articulation (left) or dynamics (right). A dashed line represents the tempo given by a metronome. Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-ioi-teaching-2}
\end{figure}

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-art-teaching-2-1} \caption{\label{fig:art-teaching-2}Experiment 2:  Comparison between those with and without piano teaching experience in terms of KOT (ms) when playing the piece with articulation. Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-art-teaching-2}
\end{figure}

\caption{\label{tab:ioi-teaching-desc-2}Experiment 2:}

\begin{tabular}{lllllllll}
\toprule
Condition & \multicolumn{1}{c}{Technique} & \multicolumn{1}{c}{Teaching} & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c}{SEM} & \multicolumn{1}{c}{Median} & \multicolumn{1}{c}{IQR}\\
\midrule
Performing & Articulation & Yes & 11 & 0.954 & 0.046 & 0.014 & 0.977 & 0.064\\
Teaching & Articulation & Yes & 11 & 0.964 & 0.049 & 0.015 & 0.976 & 0.044\\
Performing & Articulation & No & 9 & 0.951 & 0.035 & 0.012 & 0.950 & 0.041\\
Teaching & Articulation & No & 9 & 0.967 & 0.052 & 0.017 & 0.973 & 0.014\\
Performing & Dynamics & Yes & 11 & 0.953 & 0.046 & 0.014 & 0.960 & 0.050\\
Teaching & Dynamics & Yes & 11 & 0.958 & 0.047 & 0.014 & 0.972 & 0.044\\
Performing & Dynamics & No & 9 & 0.958 & 0.041 & 0.014 & 0.959 & 0.023\\
Teaching & Dynamics & No & 9 & 0.955 & 0.052 & 0.017 & 0.954 & 0.026\\
\bottomrule
\addlinespace
\end{tabular}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} Note}
\end{tablenotes}

\caption{\label{tab:art-teaching-desc-2}Experiment 2:}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} Note}
\end{tablenotes}

\caption{\label{tab:dyn-teaching-desc-2}Experiment 2:}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} Note}
\end{tablenotes}

\caption{\label{tab:dyn-diff-teaching-desc-2}Experiment 2:}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} Note}
\end{tablenotes}

>>>>

In the teaching condition, participants were instructed to play the excerpt of music as if they were teaching it to students. It was mentioned that the students already knew the sequence of the tones and that they were trying to learn how to perform the piece with the notated expression by listening to the participant's performance. In the performing condition, participants were asked to play the excerpt of music as if they were performing it to an audience (see details in \emph{\protect\hyperlink{supplementary}{Supplementary material}} 1). Participants played the piece 8 times per technique per condition, so there were 32 trials in total (2 conditions x 2 techniques x 8 trials). The order of the conditions was blocked and counterbalanced across participants. The order of the techniques within each condition was also blocked and counterbalanced across participants. A leading metronome (80 quarter beats per minute, 8 beats) indicated the target tempo before each trial.

@@

A two-way repeated-measures ANOVA with the factors Condition (teaching vs.~performing) and Dynamics (forte vs.~piano) showed that there was a significant main effect of Dynamics (\emph{F}(1,30) = 369, \emph{p} \textless{} 0.001, \(\eta_G^2\) = 0.74) and a significant interaction between Condition and Dynamics (\emph{F}(1,30) = 6.83, \emph{p} = 0.014, \(\eta_G^2\) = 0.003, \emph{Fig \ref{fig:vel-1}}). However, post-hoc comparisons based on the estimated marginal means with Tukey adjustment did not find a significant difference between the teaching {[}Forte: \emph{M} = 82.78, \emph{SD} = 8.31, Piano: \emph{M} = 60.63, \emph{SD} = 5.08{]} and the performing condition {[}Forte: \emph{M} = 81.65, \emph{SD} = 7.43, Piano: \emph{M} = 60.99, \emph{SD} = 4.41{]} for each subcomponent (Forte: \emph{p} = 0.068, Piano: \emph{p} = 0.24).

@@

We employed the same procedure as in Experiment 1 with several modifications. First, all participants were required to memorise the piece without expressive notations (i.e., \emph{Fig \ref{fig:stimuli-2} A}) prior to the experiment so that they had enough time to familiarise themselves with the piece and perform it without pitch errors while performing it with different notated expressions in the lab. Second, we modified the wording of instructions for the performing condition so that both instructions had the same focus on expressive notations because there had been less emphasis on expression in the instruction in the performing condition in Experiment 1 (see details in \emph{\protect\hyperlink{supplementary}{Supplementary material}} 1). Third, participants could choose their preferred tempo from one of three options (100, 110 and 120 quarter beats per minute) because some evidence show that musicians have their own preferred tempo, which affects their temporal variability\textsuperscript{19}. The chosen tempo was again cued by a leading metronome. To make sure that participants memorised the piece and had sufficient motor skills, we asked participants to perform the piece without looking at the original sheet music (\emph{Fig \ref{fig:stimuli-2} A}). Those who could not perform the piece without pitch errors twice consecutively within five attempts were excluded from data analysis (as a result, one participant was excluded). The rest of the procedure was identical to Experiment 1.

@@

One limitation of the study is that although we instructed participants to imagine a situation in which they were teaching musical expression to students, there was no feedback from actual students. It can be expected that teachers would also make didactic modulations in some domains other than acoustic properties, such as gestures if they are physically present in front of students. Moreover, given that teachers modulate their demonstration throughout ongoing interactions with learners\textsuperscript{20}, future studies are needed to investigate how experts dynamically adapt their performance to their students' skill levels and demonstrated abilities. Also, some of the participants in the current study did not have teaching experience at all and even those who had teaching experience had taught for only a few years. It would be important to investigate if our findings would be observed in participants with extensive experience teaching the piano. We summarised descriptive statistics depending on participants' teaching experience (see \emph{\protect\hyperlink{supplementary}{Supplementary material}} 2).

@@

We quantified participants' sound modulations in terms of tempo, articulation and dynamics so as to examine if and how participants modulate their sound. However, it is possible that participants show didactic modulations in other aspects such as temporal variability and temporal grouping (see \emph{\protect\hyperlink{supplementary}{Supplementary material}} 3 and 4). For example, participants tended to slow down at transition points regardless of the techniques they were teaching although overall, we found that participants played slower only when teaching the piece with the notated articulation. Future research is needed to investigate how musicians make particular modulations in relation to musical structures.

Musicians and music teachers tend to consider expressive skills as performers' most important skills\textsuperscript{21}. In the present study, experts focused on the teaching of expressions that were notated. However, expression also has other facets\textsuperscript{22} that are not only piece-related (e.g., notated expressions) but performer- or context-related (e.g., performer's interpretation, listening contexts such as recordings or concerts). The ultimate goal of expressive performance is considered to lie in developing performers' own interpretations of music and convey affects and emotions by using expressive tools such as articulation, dynamics, tempo and timing\textsuperscript{23}. The current research opens up the possibility of studying the teaching of expressive performance by demonstrating that pianists could signal didactic intentions without relying on verbal instructions, which are heavily used in actually teaching settings\textsuperscript{15,24}. Given that students often learn by listening to recordings\textsuperscript{25}, it is important to investigate how performance itself could affect students' learning processes.It would be interesting to examine how teachers play (i.e., demonstrate) relates to verbal instructions they give.

The present findings extend earlier research on teaching-related action modulations. First, our study sheds light on the teaching of technical skills that are an integral part of skill acquisition in artistic domains. We showed that compared to an expressive performance baseline, experts made specific modulations to teach particular techniques. Second, the specificity of the observed modulations supports the idea that teaching comprises more than generic modulations like slowing down or overall exaggeration that may draw learners' attention. Rather, expert demonstrators seem to follow principles of relevance in communication\textsuperscript{26}, highlighting only those aspects they are intending to demonstrate. How learners benefit from the perceptual and motor cues that come with specific exaggerations, and whether understanding the teacher's intentions explicitly adds to the learning success are important questions for future research.

@@

\leavevmode\vadjust pre{\hypertarget{ref-zamm_2018}{}}%
\CSLLeftMargin{19. }%
\CSLRightInline{Zamm, A., Wang, Y. \& Palmer, C. \href{https://doi.org/10.1177/0748730418783651}{Musicians' {Natural Frequencies} of {Performance Display Optimal Temporal Stability}}. \emph{Journal of Biological Rhythms} \textbf{33}, 432--440 (2018).}

\leavevmode\vadjust pre{\hypertarget{ref-okazaki_2019}{}}%
\CSLLeftMargin{20. }%
\CSLRightInline{Okazaki, S., Muraoka, Y. \& Osu, R. \href{https://doi.org/10.1038/s41598-019-44049-x}{Teacher-learner interaction quantifies scaffolding behaviour in imitation learning}. \emph{Scientific Reports} \textbf{9}, 1--13 (2019).}

@@

\leavevmode\vadjust pre{\hypertarget{ref-colprit_2000}{}}%
\CSLLeftMargin{24. }%
\CSLRightInline{Colprit, E. J. \href{https://doi.org/10.2307/3345394}{Observation and {Analysis} of {Suzuki String Teaching}}. \emph{Journal of Research in Music Education} \textbf{48}, 206--221 (2000).}

\leavevmode\vadjust pre{\hypertarget{ref-volioti_2017}{}}%
\CSLLeftMargin{25. }%
\CSLRightInline{Volioti, G. \& Williamon, A. \href{https://doi.org/10.1177/1029864916674048}{Recordings as learning and practising resources for performance: {Exploring} attitudes and behaviours of music students and professionals}. \emph{Musicae Scientiae} \textbf{21}, 499--523 (2017).}

\leavevmode\vadjust pre{\hypertarget{ref-sperber_1995}{}}%
\CSLLeftMargin{26. }%
\CSLRightInline{Sperber, D. \& Wilson, D. \emph{Relevance: {Communication} and cognition, 2nd ed}. ({Blackwell Publishing}, 1995).}

@@

\hypertarget{supplementary}{%
\section{Supplementary Material}\label{supplementary}}

@@

\hypertarget{descriptive-statistics-depending-on-participants-teaching-experience}{%
\section{2. Descriptive statistics depending on participants' teaching experience}\label{descriptive-statistics-depending-on-participants-teaching-experience}}

We summarised descriptive statistics for those had an experience in teaching the piano and those who did not. Due to the lack of statistical power, we did not perform any statistical tests.

\caption{\label{tab:ioi-teaching-desc-1}Experiment 1: Comparison between those with and without piano teaching experience in terms of IOIs (ms).}

\begin{tabular}{lllllll}
\toprule
Condition & \multicolumn{1}{c}{Technique} & \multicolumn{1}{c}{Teaching} & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c}{SEM}\\
\midrule
Performing & Articulation & Yes & 10 & 181.27 & 10.99 & 3.47\\
Teaching & Articulation & Yes & 10 & 184.72 & 13.02 & 4.12\\
Performing & Articulation & No & 21 & 193.49 & 18.75 & 4.09\\
Teaching & Articulation & No & 21 & 196.67 & 20.52 & 4.48\\
Performing & Dynamics & Yes & 10 & 182.43 & 11.52 & 3.64\\
Teaching & Dynamics & Yes & 10 & 187.15 & 11.73 & 3.71\\
Performing & Dynamics & No & 21 & 191.48 & 18.56 & 4.05\\
Teaching & Dynamics & No & 21 & 193.50 & 21.22 & 4.63\\
\bottomrule
\addlinespace
\end{tabular}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} N, SD and SEM represent sample size, standard deviation and standard error of the mean, respectively. Technique indicate performance data when participants playing the piece with either articulation (Fig.1 B) or dynamics (Fig.1 C). Teaching indicates if participants had an experience in teaching the piano.}
\end{tablenotes}

\caption{\label{tab:art-teaching-desc-1}Experiment 1: Comparison between those with and without piano teaching experience in terms of KOT (ms).}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} N, SD and SEM represent sample size, standard deviation and standard error of the mean, respectively. Technique indicate performance data when participants playing the piece with either articulation (Fig.1 B) or dynamics (Fig.1 C). Teaching indicates if participants had an experience in teaching the piano.}
\end{tablenotes}

\caption{\label{tab:dyn-teaching-desc-1}Experiment 1: Comparison between those with and without piano teaching experience in terms of KV (0-127).}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} N, SD and SEM represent sample size, standard deviation and standard error of the mean, respectively. Technique indicate performance data when participants playing the piece with either articulation (Fig.1 B) or dynamics (Fig.1 C). Teaching indicates if participants had an experience in teaching the piano.}
\end{tablenotes}

\caption{\label{tab:dyn-diff-teaching-desc-1}Experiment 1: Comparison between those with and without piano teaching experience in terms of KV Difference at transition points.}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} N, SD and SEM represent sample size, standard deviation and standard error of the mean, respectively. Technique indicate performance data when participants playing the piece with either articulation (Fig.1 B) or dynamics (Fig.1 C). Teaching indicates if participants had an experience in teaching the piano.}
\end{tablenotes}

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-ioi-teaching-2-1} \caption{\label{fig:ioi-teaching-2}Experiment 2: Comparison between those with and without piano teaching experience in terms of normalised IOIs when playing the piece with either articulation (left) or dynamics (right). A dashed line represents the tempo given by a metronome. Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-ioi-teaching-2}
\end{figure}

\begin{figure}
\includegraphics[width=1\linewidth]{manuscript_files/figure-latex/plot-art-teaching-2-1} \caption{\label{fig:art-teaching-2}Experiment 2:  Comparison between those with and without piano teaching experience in terms of KOR when playing the piece with articulation. Each box indicates the IQR with the median, and whiskers extend to a maximum of 1.5 × IQR beyond the box.}\label{fig:plot-art-teaching-2}
\end{figure}

\caption{\label{tab:ioi-teaching-desc-2}Experiment 2: Comparison between those with and without piano teaching experience in terms of normalised IOIs.}

\begin{tabular}{lllllll}
\toprule
Condition & \multicolumn{1}{c}{Technique} & \multicolumn{1}{c}{Teaching} & \multicolumn{1}{c}{N} & \multicolumn{1}{c}{Mean} & \multicolumn{1}{c}{SD} & \multicolumn{1}{c}{SEM}\\
\midrule
Performing & Articulation & Yes & 11 & 0.954 & 0.046 & 0.014\\
Teaching & Articulation & Yes & 11 & 0.964 & 0.049 & 0.015\\
Performing & Articulation & No & 9 & 0.951 & 0.035 & 0.012\\
Teaching & Articulation & No & 9 & 0.967 & 0.052 & 0.017\\
Performing & Dynamics & Yes & 11 & 0.953 & 0.046 & 0.014\\
Teaching & Dynamics & Yes & 11 & 0.958 & 0.047 & 0.014\\
Performing & Dynamics & No & 9 & 0.958 & 0.041 & 0.014\\
Teaching & Dynamics & No & 9 & 0.955 & 0.052 & 0.017\\
\bottomrule
\addlinespace
\end{tabular}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} N, SD and SEM represent sample size, standard deviation and standard error of the mean, respectively. Technique indicate performance data when participants playing the piece with either articulation (Fig.5 B) or dynamics (Fig.5 C). Teaching indicates if participants had an experience in teaching the piano.}
\end{tablenotes}

\caption{\label{tab:art-teaching-desc-2}Experiment 2: Comparison between those with and without piano teaching experience in terms of KOR.}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} N, SD and SEM represent sample size, standard deviation and standard error of the mean, respectively. Technique indicate performance data when participants playing the piece with either articulation (Fig.5 B) or dynamics (Fig.5 C). Teaching indicates if participants had an experience in teaching the piano.}
\end{tablenotes}

\caption{\label{tab:dyn-teaching-desc-2}Experiment 2: Comparison between those with and without piano teaching experience in terms of KV (0-127).}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} N, SD and SEM represent sample size, standard deviation and standard error of the mean, respectively. Technique indicate performance data when participants playing the piece with either articulation (Fig.5 B) or dynamics (Fig.5 C). Teaching indicates if participants had an experience in teaching the piano.}
\end{tablenotes}

\caption{\label{tab:dyn-diff-teaching-desc-2}Experiment 2: Comparison between those with and without piano teaching experience in terms of KV Difference at transition points.}

\begin{tablenotes}[para]
\normalsize{\textit{Note.} N, SD and SEM represent sample size, standard deviation and standard error of the mean, respectively. Technique indicate performance data when participants playing the piece with either articulation (Fig.5 B) or dynamics (Fig.5 C). Teaching indicates if participants had an experience in teaching the piano.}
\end{tablenotes}

\hypertarget{temporal-variability}{%
\section{3. Temporal Variability}\label{temporal-variability}}

@@

\hypertarget{temporal-grouping}{%
\section{4. Temporal grouping}\label{temporal-grouping}}

@@

\hypertarget{additional-analysis-for-dynamics-difference}{%
\section{5. Additional analysis for dynamics difference}\label{additional-analysis-for-dynamics-difference}}

@@
